---
title: "Homework 5"
author: Troy Layouni
output: github_document
---

```{r setup, include=FALSE}
library(viridis)
library(tidyverse)
library(readr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

**Loading in `iris` dataset with missing values**

```{r load_iris_with_missing}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

#### Writing a function to replace missing values 

* missing numeric values are filled in with the mean of non-missing values 
* missing character values are filled in with `"virginica"

```{r no_missing_df}
fix_missing = function(x) {
  if (is.numeric(x)) {
    replace_na(x, mean(x, na.rm = TRUE))
  } else if (is.character(x)) {
  replace_na(x, "virginica")
  }
}

iris_no_missing = map_dfr(iris_with_missing, fix_missing)
```

## Problem 2

**Loading and tidying control and experimental arm data** 

* reading in all files from each subject into one dataframe
* tidying data to include 4 columns with study arm, id, week and 


```{r study_data_load}
read_files = function(file) {
  study_data = read_csv(file)
}

path_base = "./data/"

join_data = 
  tibble(
    file_name = list.files(path = "./data/"),
    all_files = str_c(path_base, file_name)) %>% 
  mutate(study_data = map(all_files, read_files)) %>% 
  unnest() %>% 
  select(-all_files) %>% 
  separate(file_name, into = c("study_arm", "subject_id"), sep = "\\_") %>% 
  pivot_longer(
    week_1:week_8, 
    names_to = "week", 
    values_to = "observation"
    ) %>% 
  mutate(
    subject_id = str_replace(subject_id, ".csv", ""),
    week = str_replace(week, "week_", ""),
    study_arm = str_replace(study_arm, "con", "control"),
    study_arm = str_replace(study_arm, "exp", "experimental")
)
```

